HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2023-05-30 tmlr_submission/hermes_paper_revision/hermes_paper.tex
Update hermes_paper.tex

Following this work, we present in this paper HERMES, a new hybrid recurrent model for time series forecasting with inclusion of external signals. This new architecture is decomposed  into two parts: \textcolor{gray}{local predictors and a global corrector. First, a per-time-series parametric statistical model is trained on each sequence. Then, a global recurrent neural network is trained to evaluate and correct the forecast weaknesses of the first collection of models.} \textcolor{orange}{i) First, a per-time-series parametric statistical model is trained on each sequence. ii) Then, a global recurrent neural network is trained to evaluate and correct the forecast weaknesses of the first collection of models.} The external weak signals reveal the real potential of the hybrid approach: a global neural network, able to leverage large amounts of heterogeneous data, deal with any kind of external weak signals, learn context and finally correct weaknesses and errors of parametric models.

@@

During the M4 competition, the hybrid model of \citet{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. \textcolor{orange}{Thus, the choice of the parametric model can be adjusted depending on the nature of the time series. The only limitation is the computation time. For instance, hidden Markov models or ARIMA models are powerfull parametric methodes but they are also computational expensive to fit. Choosing these approaches as per-time-series predictor may considerably slow the overall training process of the hybrid model. In Section~\ref{sec:exp}, three declinations of the hybrid framework are presented using different per-time-series predictors. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020}, the second one use a method called Thetam as per-time-series predictor \citep{hyndman2020} and the last one uses a TBATS model \citep{alysha2011}.} \textcolor{gray}{In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020} and the second one uses a TBATS model \citep{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.}

@@

As benchmarks, several widespread statistical methods and deep learning approaches were selected. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats}, for each time series, predictions are computed with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast and \textit{tbats} uses a trigonometrical seasonality. Finally, \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citet{hyndman2020}. \textcolor{orange}{To complete the statistical benchmarks, the Prophet model introduced in \citet{Taylor2017} is proposed and implemented with the available package bearing the same name \texttt{prophet}. As deep learning apporaches, Three models based on recurrent neural networks are proposed. They are called \textit{lstm}, \textit{lstm-ws} and \textit{deepar}. Concerning \textit{lstm} and \textit{lstm-ws}, they are both composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52. The first one (\textit{lstm}) has only access to the main signal while the second one (\textit{lstm-ws}) has access to the main and the external signal. The last approach is the recent state-of-the-art model called DeepAR and introduced in \citet{salinas2020}. The python package \texttt{GluonTS} \citep{Alexandrov2020} was used to train the DeepAR approach on the fashion dataset and a grisearch on the main parameters have been done.} \textcolor{gray}{As a deep learning approach, a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52 is considered.}

\textcolor{orange}{
A strength of the proposed HERMES framework is that it can handle any kind of parametric model. This choice of per-time-series predictor is central in the approach and have to be done regarding the use case and the nature of the time series. Thus, three declinations of the HERMES approach are proposed on the Fashion dataset using different per-time-series predictors: Ets, Tbats and Thetam. They are called respectively \textit{hermes-ets}, \textit{hermes-tbats} and \textit{hermes-thetam} according to the per-time-series model choice. For each of them, a variation with the inclusion of the weak signals (\textit{-ws}) is presented.}
\textcolor{gray}{Two versions of HERMES are proposed. They are called respectively \textit{hermes-ets} and \textit{hermes-tbats} according to the per-time-series model choice. Moreover, two versions with the inclusion of the weak signals (ws) are proposed. They are referred to as \textit{hermes-ets-ws} and \textit{hermes-tbats-ws}. In order to provide a fair comparison, a \textit{lstm} with the weak signals named \textit{lstm-ws} is trained.}

@@

\textcolor{orange}{Considering the new HERMES approaches, all the different versions show consequent improvements regarding their per-time-series predictors in terms of MASE. \textit{Hermes-tbats} largely outperforms the two other declinations and all the benchmarks on the fashion dataset. This result highlight to important point of the proposed hybrid model: i) The final accuracy of the model is strongly impacted by the choice of per-time-series predictor ii) If the per-time-series predictor is well chosen regarding the nature of the time series, the resulting HERMES model can show impressive results and outperformance state-of-the-art approaches such as DeepAR.}
\textcolor{gray}{Considering the new HERMES approach, \textit{hermes-tbats} and \textit{hermes-tbats-ws} slightly outperform the alternatives in terms of MASE and are stable across the different trainings. Regarding \textit{hermes-ets},   although it is very similar to the baseline \citet{smyl2020}, its accuracy remains low in comparison to the \textit{lstm} benchmark or HERMES using TBATS.}

\textcolor{orange}{Regarding the impact of the weak signals, improvement in terms of MASE remains low between methodes without and with the weak signal. However, an interesting improvement can be noted with the accuracy metric. Weak signals seem to grandtully help methods using them to better detect emerging and declining trends. Figure~\ref{fig:examples} displays some examples of \textit{hermes-tbats} predictions and illustrates some Tbats weaknesses that can be corrected by the hybrid approach.}
\textcolor{gray}{
Models using our weak signals perform similarly as without-weak-signals models for the MASE.  Interestingly, weak signals significantly improve the accuracy in detecting emerging and declining trends. Figure~\ref{fig:examples} displays some examples of \textit{hermes-tbats} models and some weaknesses that can be corrected.}

@@

\textbf{Size of the dataset. } In addition to the results on the whole fashion dataset, the robustness of the HERMES model is analyzed when it is trained on smaller datasets. Two experiments are performed on a sub sample of respectively 1000 and 100 randomly selected time series. Results are given in Table~\ref{tab:1000metricresults}. The hybrid framework \textit{hermes-tbats} achieves the best performance in terms of global accuracy on both datasets. We can note that the accuracy of the full neural network \textit{lstm} decreases when the dataset size decreases. On the small dataset of 100 time series, a local statistical model like \textit{tbats} or \textit{stlm} largely outperforms \textcolor{orange}{deep-model-based approach such as \textit{lstm} of \textit{deepar}}. In fact, providing sharp predictions from scratch is a complex task and high-dimensional recurrent neural networks require large amounts of data to do so. By contrast, the HERMES approach can rely on its first statistical part and consequently needs less data to be trained and to obtain interesting performance. \textcolor{orange}{We can nevertheless note that the gain brought by the error-corrector recurrent model decreases as the size of the dataset diminish.}

>>>>

Following this work, we present in this paper HERMES, a new hybrid recurrent model for time series forecasting with inclusion of external signals. This new architecture is decomposed  into two parts: \textcolor{gray}{local predictors and a global corrector. First, a per-time-series parametric statistical model is trained on each sequence. Then, a global recurrent neural network is trained to evaluate and correct the forecast weaknesses of the first collection of models.} \textcolor{orange}{First, a per-time-series parametric statistical model is trained on each sequence. Then, a global recurrent neural network is trained to evaluate and correct the forecast of the first collection of models.} The external weak signals reveal the real potential of the hybrid approach: a global neural network, able to leverage large amounts of heterogeneous data, deal with any kind of external weak signals, learn context and finally correct weaknesses and errors of parametric models.

@@

During the M4 competition, the hybrid model of \citet{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. \textcolor{orange}{Thus, the choice of the parametric model can be adjusted depending on the nature of the time series. The only limitation is the computation time as we aim at forecasting thousands of time series simultaneously. For instance, hidden Markov models provide a very appealing framework but inference of such models requires computationally costly iterative procedures such as Expecation Maximization-based algorithms often combined with Monte Carlo estimates of unknown expectations. Choosing these approaches as per-time-series predictor would considerably slow the overall training process of the hybrid model.}

\textcolor{orange}{In Section~\ref{sec:exp}, we introduce three declinations of the hybrid framework  using different per-time-series predictors to highlight the adaptability of our approach. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020}, the second one use a method called Thetam as per-time-series predictor \citep{hyndman2020} and the last one uses a TBATS model \citep{alysha2011}.} \textcolor{gray}{In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020} and the second one uses a TBATS model \citep{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.}

@@

As benchmarks, several widespread statistical methods and deep learning approaches were selected. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats}, for each time series, predictions are computed with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast and \textit{tbats} uses a trigonometrical seasonality. Finally, \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citet{hyndman2020}.

\textcolor{orange}{We also compare our model with recent recent deep learning architectures for time series. The Prophet model introduced in \citet{Taylor2017} is proposed and implemented with the available package bearing the same name \texttt{prophet}, and three models based on recurrent neural networks called \textit{lstm}, \textit{lstm-ws} and \textit{deepar} are used. Concerning \textit{lstm} and \textit{lstm-ws}, they are both composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52. The first one (\textit{lstm}) has only access to the main signal while the second one (\textit{lstm-ws}) has also access to the external signal. The last approach is the recent state-of-the-art model called DeepAR and introduced in \citet{salinas2020}. The Python package \texttt{GluonTS} \citep{Alexandrov2020} was used to train DeepAR on the fashion dataset with a gridsearch to tune the main hyperparameters.} \textcolor{gray}{As a deep learning approach, a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52 is considered.}

\textcolor{orange}{
A strength of the proposed HERMES framework is that it can handle any kind of parametric model. Thus, three declinations of the HERMES approach are proposed on the Fashion dataset using different per-time-series predictors: Ets, Tbats and Thetam. They are called respectively \textit{hermes-ets}, \textit{hermes-tbats} and \textit{hermes-thetam} according to the per-time-series model choice. For each of them, a variation with the inclusion of the weak signals (\textit{-ws}) is presented.}
\textcolor{gray}{Two versions of HERMES are proposed. They are called respectively \textit{hermes-ets} and \textit{hermes-tbats} according to the per-time-series model choice. Moreover, two versions with the inclusion of the weak signals (ws) are proposed. They are referred to as \textit{hermes-ets-ws} and \textit{hermes-tbats-ws}. In order to provide a fair comparison, a \textit{lstm} with the weak signals named \textit{lstm-ws} is trained.}

@@

\textcolor{orange}{All the different versions of HERMES show consequent improvements regarding their per-time-series predictors in terms of MASE. \textit{Hermes-tbats} largely outperforms the two other declinations and all the benchmarks on the fashion dataset. This result highlights two important features of the proposed hybrid model: i) the final accuracy of the model is strongly impacted by the choice of the per-time-series predictor and ii) if the per-time-series predictor is well chosen regarding the nature of the time series, the resulting HERMES model outperforms state-of-the-art approaches such as DeepAR.}
\textcolor{gray}{Considering the new HERMES approach, \textit{hermes-tbats} and \textit{hermes-tbats-ws} slightly outperform the alternatives in terms of MASE and are stable across the different trainings. Regarding \textit{hermes-ets},   although it is very similar to the baseline \citet{smyl2020}, its accuracy remains low in comparison to the \textit{lstm} benchmark or HERMES using TBATS.}

\textcolor{orange}{Regarding the impact of the weak signals, Table~\ref{tab:metricresults} highlights an interesting improvement of the accuracy metric when weak signals are included. Figure~\ref{fig:examples} displays some examples of \textit{hermes-tbats} predictions and illustrates some Tbats weaknesses that can be corrected by the hybrid approach.}
\textcolor{gray}{
Models using our weak signals perform similarly as without-weak-signals models for the MASE.  Interestingly, weak signals significantly improve the accuracy in detecting emerging and declining trends. Figure~\ref{fig:examples} displays some examples of \textit{hermes-tbats} models and some weaknesses that can be corrected.}

@@

\textbf{Size of the dataset. } In addition to the results on the whole fashion dataset, the robustness of the HERMES model is analyzed when it is trained on smaller datasets. Two experiments are performed on a sub sample of respectively 1000 and 100 randomly selected time series. Results are given in Table~\ref{tab:1000metricresults}. The hybrid framework \textit{hermes-tbats} achieves the best performance in terms of global accuracy on both datasets. We can note that the accuracy of the full neural network \textit{lstm} decreases when the dataset size decreases. On the small dataset of 100 time series, a local statistical model like \textit{tbats} or \textit{stlm} largely outperforms \textcolor{orange}{deep-model-based approach such as \textit{lstm} of \textit{deepar}}. In fact, providing sharp predictions from scratch is a complex task and high-dimensional recurrent neural networks require large amounts of data to do so. By contrast, the HERMES approach can rely on its first statistical part and consequently needs less data to be trained and to obtain interesting performance. \textcolor{orange}{We can nevertheless note that the gain brought by the error-corrector recurrent model decreases as the size of the dataset diminishes.}
